Static Race Detection in OS Kernels by Mining Locking Rules

### 2023-07-27 4.framework.tex
Finish the draft of technique
Finish the draft of framework

% !TeX spellcheck = en_US
\section{Framework}
\label{sec_framework}

>>>>

% !TeX spellcheck = en_US
\section{\sys Approach}
\label{sec_framework}

Based on the three key techniques in Section~\ref{sec_technique}, we develop a 
practical static approach named \sys, to detect data races in OS kernels, and 
estimate harmfulness of these data races. We have implemented \sys with 
Clang~\cite{Clang} and Z3~\cite{Z3}. \sys automatically mines locking rules, 
detects data races and estimates the harmfulness of detected data races. 
Figure~\ref{fig_architecture} shows the architecture of \sys, which has four 
phases:

\begin{figure}[htbp]
	\centering
	\includegraphics[width=1\linewidth]{figures/fig_architecture.pdf}
	\figcaption{\sys architecture.}
	\label{fig_architecture}
\end{figure}

\PP{P1: Source-code compilation.} The Clang compiler compiles the OS source 
code into LLVM bytecode, and then the information collector scans each LLVM 
bytecode file to record function information (including the position of each 
function definition and function name, etc.) in a database. Such information is 
used in subsequent code analysis for inter-procedural analysis across source 
files.

\PP{P2: Locking-rule Mining.} The locking-rule miner uses our alias-aware rule 
mining method to deduce locking rules about whether a variable access should be 
protected by a lock and which lock is required.

\PP{P3: Data race detection and false-positive filtering.} The data race 
detector detecting data races by checking whether a given access violate the 
rules mined by our locking-rule miner. If so, the data race filter drops 
false-positives with our locking-usage analysis. Besides, for a given data 
race, there may be multiple code paths from the entry function to its 
problematic instruction, and thus many repeated data races can be reported.  
To drop repeated data races, for a new possible data race, the data race filter 
checks whether its problematic instruction is identical to any existing data 
race. If so, this possible data race is regarded as repeated and dropped.

\PP{P4: Harmfulness estimation.} The harmfulness estimator uses our 
pattern-based estimation to estimate the harmfulness of detected data races, 
and reports data races that can cause memory or logic bugs. With these result, 
developers can focus on those harmful data races.
